Versatile Robust Clustering of Ad Hoc Cognitive Radio Network

### 2016-12-10 journal-draft/bare_jrnl_comsoc.tex
minor modification, before full polishing

There has been a lot of research done for clustering in wireless networks.
In ad-hoc and mesh networks, the major goal of clustering is to maximize connectivity or to improve the performance of routes~\cite{Kawadia03, clustering_mesh_globecom2010}.
The emphasis of clustering in sensor networks is on network lifetime and coverage~\cite{Abbasi_survey_07}.
%Various clustering schemes are proposed to target different aspects in cognitive radio networks.
%Work~\cite{Consensus_based_clustering12} improves spectrum sensing ability by grouping the CR users with potentially best detection performance into the same cluster.
%Clustering scheme~\cite{clustering_globecom11} obtains the best cluster size which minimizes power consumption caused by communication within and among clusters.
%\cite{clustering_globecom11} proposes clustering strategy in cognitive radio network, which looks into the relationship between cluster size and power consumption and accordingly controlling the cluster size to decrease power consumption.
In respect of CRN, \cite{Zhao07, Chen07,Affinity_clustering_09icccn} propose the clustering schemes to form clusters, where securing CCs is the only goal.
Clustering scheme~\cite{Consensus_based_clustering12} improves spectrum sensing accuracy with cluster structure.
%Clustering scheme~\cite{clustering_globecom11} obtains the best cluster size which minimizes power consumption caused by communication within and among clusters.
\cite{clustering_globecom11, TWC2012_cooperative_communication} target on the QoS provisioning and energy efficiency with cluster structure.
\cite{cluster_EW10} forms clusters to coordinate the control channel usage within one cluster.
An event-driven clustering scheme is proposed for cognitive radio sensor network in \cite{Ozger_cluster_crsn_13}.
No one among the above mentioned schemes provides the robustness to the formed clusters against primary users. 
%
A clustering scheme (denoted as SOC) which is designed to generate robust clusters against primary users is proposed in~\cite{LIU_TMC11_2}.
SOC involves three phases of distributed executions.
In the first phase, every secondary user forms clusters with some one-hop neighbors, in the second and third phase, each secondary user seeks to either merge other clusters or join one of them.
The metric adopted by every secondary user in the three phases is the product of the number of CCs and cluster size.
The drawbacks are as follows, although the adopted metric considers both cluster size and the number of CCs, cluster formation can be easily dominated by only one factor, e.g. a node which can access many channels may exclude its neighbor and form a cluster by itself.
In addition, this scheme leads to the high variance of the cluster sizes, which is not desired in certain applications as discussed in~\cite{clustering_globecom11, cluster_EW10}.
\cite{mansoor_15_cluster_robust} presents a heuristic method to form clusters, although the authors claim robustness is one goal to achieve, the minimum number of clusters is finally pursued.

A distributed clustering scheme ROSS is proposed in~\cite{Li11_ROSS} under the game theoretic framework. 
Compared with the clustering schemes introduced above, the clusters are formed faster and the clusters possess more CCs within and among clusters than SOC.
But this work doesn't intervene the outcome of singleton clusters, although the cluster sizes are not as divergent as SOC.
Furthermore, this work doesn't consider the robustness of clusters against the increasing activity of primary users, which leaves their claim of robustness unverified.
%
This paper is on the basis of the work in~\cite{Li11_ROSS}, but extends in two directions.
First, this paper sticks to a new metric of robustness, and considers the clusters when primary users' activity becomes more dynamic.
Second, this paper proposes size control mechanism, which solves the problem of devergence of clusters sizes in ~\cite{Li11_ROSS} and~\cite{LIU_TMC11_2}.
Besides, this paper provides a comprehensive analysis of the robust clustering problem and proposes the centralized solution. 
%We stick to the motivation of forming robust clusters in CRN \ie let more CR users benefited from the cooperative decision making which is due to the clusters. 
%We propose both centralized and distributed clustering schemes, which result in more CR users in the clusters composed with multiple CR users, not only when the clusters are just formed, but also afterwards when the primary users' activity changes.
%Besides, both centralized and distributed schemes take the cluster size into consideration, and both can control the sizes of the formed clusters to certain extent.
%In particular, the decentralized clustering approach ROSS (RObust Spectrum Sharing) is able to form clusters with desired sizes.
%When compared with other distributed works, ROSS involves smaller signaling overhead and the generated clusters are significantly more robust against the primary users which appear after the clusters are formed.
%thesis
%and the generated clusters are more robust than the other clustering scheme which has claims on cluster robustness, \ie more secondary users residing in clusters against increasing influence from primary users.
%
%ROSS selects cluster head through coordination within its neighborhood, and then cluster membership is decided locally and its fast convergence is proved under game theoretic framework. 
%For our scheme we can prove convergence in cluster formation phase and resolve ambiguities with respect to cluster membership in a game-theoretic setting. 
The new extensions are made on basis of ROSS and its light weighted version, the latter involves less overheads thus are more suitable for the scenario where fast deployment is desired.
Throughout this paper, we refer to the clustering schemes on the basis of ROSS as \textit{variants of ROSS}, which include the fast versions and that with size control function.

@@

Given a cognitive radio network $\mathcal{N}$ where $|\mathcal{N}|=N$, the collection of all the possible clusters\footnote{Possible cluster means the collection of CR nodes, which complies with the definition in~\ref{sec:model}} in $\mathcal{N}$ is denoted as $\mathcal{S}$ where $\mathcal{S}=\{C_1, C_2,\ldots,C_{|\mathcal{S}|}\}$ \footnote{The subscripts of the clusters can be decided in any convenient way, \ie the sequence of identifying them.} and there is $\bigcup_{1\leq i \leq |\mathcal{S}|} C_i = N$.
%
With the requirements on the cluster size are enforced, \ie the desired size is $\delta$ and the cluster sizes should fall in the scope $\langle\delta_1, \delta_2\rangle$, where $\delta, \delta_1, \delta_2\in \mathbb{Z}^+$, and $\delta_1 \leq \delta\leq \delta_2$, a feasible clustering solution is a subcollection $\mathcal{S}' \subseteq \mathcal{S}$, which satisfies 
$\delta_1\leq|C_i|\leq \delta_2, \bigcup_{C_i\in \mathcal{S}'} C_i = \mathcal{N}$ and $C_{i'}\cap C_i =\emptyset$ where $C_{i'}, C_i\in \mathcal{S}'$ and $i'\neq i$.
The optimal clustering solution is the feasible clustering solution whose sum of the numbers of CCs of clusters $\sum_{C\in \mathcal{S}'} f(C)$ is the maximal.

>>>>

There has been a lot of research done for clustering in wireless networks.
In ad-hoc and mesh networks, the major goal of clustering is to maximize connectivity or to improve the performance of routes~\cite{Kawadia03, clustering_mesh_globecom2010}.
The emphasis of clustering in sensor networks is on network lifetime and coverage~\cite{Abbasi_survey_07}.
%Various clustering schemes are proposed to target different aspects in cognitive radio networks.
%Work~\cite{Consensus_based_clustering12} improves spectrum sensing ability by grouping the CR users with potentially best detection performance into the same cluster.
%Clustering scheme~\cite{clustering_globecom11} obtains the best cluster size which minimizes power consumption caused by communication within and among clusters.
%\cite{clustering_globecom11} proposes clustering strategy in cognitive radio network, which looks into the relationship between cluster size and power consumption and accordingly controlling the cluster size to decrease power consumption.
In respect of CRN, \cite{Zhao07, Chen07,Affinity_clustering_09icccn} propose the clustering schemes to form clusters, where securing CCs is the only goal.
Clustering scheme~\cite{Consensus_based_clustering12} improves spectrum sensing accuracy with cluster structure.
%Clustering scheme~\cite{clustering_globecom11} obtains the best cluster size which minimizes power consumption caused by communication within and among clusters.
\cite{clustering_globecom11, TWC2012_cooperative_communication} target on the QoS provisioning and energy efficiency with cluster structure.
\cite{cluster_EW10} forms clusters to coordinate the control channel usage within one cluster.
An event-driven clustering scheme is proposed for cognitive radio sensor network in \cite{Ozger_cluster_crsn_13}.
No one among the above mentioned schemes provides the robustness to the formed clusters against primary users. 
%
A distributed clustering scheme (denoted as SOC) which is designed to generate robust clusters against primary users is proposed in~\cite{LIU_TMC11_2}.
The robustness of the clusters is obtained from the multiple CCs.
SOC involves three phases of distributed executions.
In the first phase, every secondary user forms clusters with some one-hop neighbors, in the second and third phase, each secondary user seeks to either merge other clusters or join one of them.
The metric adopted by every secondary user in all the phases is the product of the number of CCs and cluster size.
The drawbacks are as follows, although the adopted metric considers both cluster size and the number of CCs, cluster formation can be easily dominated by only one factor, e.g. a node which can access many channels may exclude its neighbor and form a cluster by itself.
In addition, this scheme leads to the high variance of the cluster sizes, which is not desired in certain applications as discussed in~\cite{clustering_globecom11, cluster_EW10}.
\cite{mansoor_15_cluster_robust} presents a heuristic method to form clusters, although the authors claim robustness is one goal to achieve, the minimum number of clusters is finally pursued.

A distributed clustering scheme ROSS is proposed in~\cite{Li11_ROSS} under the game theoretic framework. 
Compared with the clustering schemes introduced above, the clusters are formed faster and the clusters possess more CCs within and among clusters than SOC.
But as all the other clustering schemes, this work doesn't intervene the outcome of the very small or very large clusters which are not desirable, although the cluster sizes are not as divergent as SOC.
Furthermore, this work doesn't consider the robustness of clusters against the increasing activity of primary users, which leaves their claim of robustness unverified.
%
This paper is on the basis of the work in~\cite{Li11_ROSS}, but extends in two directions.
First, this paper sticks to a new metric of robustness, and considers the clusters when primary users' activity becomes more dynamic.
Second, this paper proposes size control mechanism, which solves the problem of devergence of clusters sizes in ~\cite{Li11_ROSS} and~\cite{LIU_TMC11_2}.
Besides, this paper provides a comprehensive analysis of the robust clustering problem and proposes the centralized solution. 
%We stick to the motivation of forming robust clusters in CRN \ie let more CR users benefited from the cooperative decision making which is due to the clusters. 
%We propose both centralized and distributed clustering schemes, which result in more CR users in the clusters composed with multiple CR users, not only when the clusters are just formed, but also afterwards when the primary users' activity changes.
%Besides, both centralized and distributed schemes take the cluster size into consideration, and both can control the sizes of the formed clusters to certain extent.
%In particular, the decentralized clustering approach ROSS (RObust Spectrum Sharing) is able to form clusters with desired sizes.
%When compared with other distributed works, ROSS involves smaller signaling overhead and the generated clusters are significantly more robust against the primary users which appear after the clusters are formed.
%thesis
%and the generated clusters are more robust than the other clustering scheme which has claims on cluster robustness, \ie more secondary users residing in clusters against increasing influence from primary users.
%
%ROSS selects cluster head through coordination within its neighborhood, and then cluster membership is decided locally and its fast convergence is proved under game theoretic framework. 
%For our scheme we can prove convergence in cluster formation phase and resolve ambiguities with respect to cluster membership in a game-theoretic setting. 
The new extensions are made on basis of ROSS and its light weighted version, the latter involves less overheads thus are more suitable for the scenario where fast deployment is desired.
Throughout this paper, we refer to the clustering schemes on the basis of ROSS as \textit{variants of ROSS}, which include the fast versions and that with size control function.

@@

Given a cognitive radio network $\mathcal{N}$ where $|\mathcal{N}|=N$, the collection of all the possible clusters\footnote{Possible cluster means the collection of CR nodes, which complies with the definition in~\ref{sec:model}} in $\mathcal{N}$ is denoted as $\mathcal{S}$ where $\mathcal{S}=\{C_1, C_2,\ldots,C_{|\mathcal{S}|}\}$ \footnote{The subscripts of the clusters can be decided in any convenient way, \ie the sequence of identifying them.} and there is $\bigcup_{1\leq i \leq |\mathcal{S}|} C_i = N$.
%
When the requirements on the cluster size are enforced, \ie the desired size is $\delta$ and the cluster sizes should fall in the scope $\langle\delta_1, \delta_2\rangle$, where $\delta, \delta_1, \delta_2\in \mathbb{Z}^+$, and $\delta_1 \leq \delta\leq \delta_2$, a feasible clustering solution is a subcollection $\mathcal{S}' \subseteq \mathcal{S}$, which satisfies 
$\delta_1\leq|C_i|\leq \delta_2, \bigcup_{C_i\in \mathcal{S}'} C_i = \mathcal{N}$ and $C_{i'}\cap C_i =\emptyset$ where $C_{i'}, C_i\in \mathcal{S}'$, $i'\neq i$.

The robust clustering problem in CRN is to find the optimal clustering solution from the feasible clustering solutions, whose sum of the numbers of CCs of clusters $\sum_{C\in \mathcal{S}'} f(C)$ is the maximal.
